\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2018-11-14 draft.tex
polish

MapReduce framework for data-parallel computation was first proposed by Google~\cite{dean04} and later implemented in the Apache Hadoop project~\cite{hadoop}.
Under the MapReduce framework, the reducer computes output values from a sequence of input values transmitted over the network.  Due to the non-determinism in data transmission, the order of input values arrived at the reducer is not fixed.
The \emph{commutativity problem} of reducers asks if the output of a reducer is independent of the order of its inputs. There are several advantages for a reducer being commutative, e.g., the verification problem of a MapReduce program can be reduced to the problem of verifying a sequential program. 
In this paper, we propose effective heuristics for reducer commutative analysis and implement them as a tool J-ReCoVer (Java Reducer Commutativity Verifier), which is the first tool that is specialized in checking reducer commutativity. Experimental results over 128 benchmark examples collected from literature and open repositories are very positive; J-ReCoVer correctly handles over 97\% of them.

@@

On the other hand, in many cases, the non-commutative behavior of a reducer can be the source of very tricky bugs. A study conducted by Microsoft investigated the commutativity problem of $508$ reducers running on their MapReduce server~\cite{xiao14mr}. The reducer programs have been checked carefully using all traditional means such as code review, testing, and experiments with real data for months. Still, among them, five programs contain very subtle bugs caused by non-commutativity (confirmed by the programmers).

@@

In this paper, we propose heuristics for the commutativity problem that works very efficiently on a large set of practical integer reducer programs. We reduce the commutativity problem to a SMT problem in a sound but incomplete manner. We also propose optimizations to enable the approach working on real world examples. For the case that the reducer is not proven commutative, we complement the approach by using testing to find concrete counterexamples. We implement those heuristics as a tool called J-ReCoVer (Java Reducer Commutativity Verifier), which is available at~\verb|http://jrecover.iis.sinica.edu.tw|.

@@

One main motivation of commutativity analysis is to reduce the verification problem of a concurrent program to a sequential one. There are several researches sharing the similar motivation. For instance, the concept of \emph{robustness} of event-driven asynchronous programs~\cite{ahmed2017:robustness}  and program running under relaxed memory model~\cite{ahmed2013:robustness,AbdullaACLR13,AbdullaACLR12}, \emph{synchronizability} of communicating finite state machines~\cite{FinkelL17,basu2012synchronizability,basu2011choreography}. However, their computation models are quite different to that of reducers and hence the result cannot be directly applied. Besides verification, another interesting research direction is the synthesis of MapReduce programs~\cite{SmithA16}. Commutativity analysis is used as a component under the synthesis framework.

\section{Notations and Definitions}
\label{section:integer-reducers}
We use $\Z$ to denote the set of all integers and $[n,m]$ to denote the set of integers $\{k\mid n \leq k\leq m\}$. 
We lift the equality $=$ operation to tuples in the standard way.
Let $S\subseteq [1,k]$, the equality $\{v_1,\ldots,v_k\}=_{S} \{v'_1,\ldots,v'_k\}$ between tuples holds if and only if $\bigwedge_{i\in S} v_i=v'_i$.
For presentation purpose, we describe a highly simplified model to characterize the essence of reducers. The model allows to describe the computation of reducers and investigate their commutativity.

\begin{equation*}
\begin{array}{rclcl}
\multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and Constant}\\
e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
g\in \Grd  & {\triangleq}  & e < e \mid not\ g \mid g \wedge g  \mid \ldots& & \textmd{Guard}\\
s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && \textmd{Command}\\
p\in \Prg  & {\triangleq}  & s.\rloop\{s\}.s&& \textmd{Reducer Program}\\
\end{array}
\end{equation*}

Here $v\in \Var$ and $c \in \Z$ are variables and constant integer values, respectively. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{guard} in $\Grd$ can be a comparison between two expressions or boolean combinations of several guards. A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each branch\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing every branch statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g)\wedge s_1$ or $(not\ g) \wedge s_2$.}. According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. A few examples of reducer programs in this model can be found in Figure~\ref{fig:reducer_example}.

@@

In the next section, we give a brief overview of J-Recover,  our tool for reducer commutativity verification.

\section{Overview of the J-ReCoVer tool}
\label{sec:overview}
%Hadoop, Java;
The input of J-ReCoVer is a reducer program written in Java, which is the most popular programming language used in the Hadoop~\cite{hadoop} MapReduce framework. 
The J-ReCoVer tool has three main components, the \emph{Preprocessor}, the \emph{Prover}, and the \emph{BugFinder}. As the name suggests, the Preprocessor reads as input a reducer program and performs the required preprocessing.  The goal of Prover is to show that a given reducer is commutative and the goal of BugFinder is the opposite. The architecture of J-ReCoVer can be found in Figure~\ref{fig:overview-jrecover}. The user can input a reducer program to J-ReCoVer either through our web-interface or binary installed on their own machine.

@@

\textbf{The Preprocessor component} first compiles a reducer program to bytecode and use the tool Soot~\cite{soot} to further convert it to the so called ``Jimple'' format, which is an intermediate language that is designed to simplify the analysis of Java programs. Under the Hadoop MapReduce framework, the permutation of the input is handled by the scheduler/shuffler component and is affected by things like network latency, which is not controllable by programmers. In order to deal with this issue, we wrote our own dummy Hadoop environment for the reducer as a part of the Preprocessor component. So that the input order of the reducer is now controlled by J-ReCoVer. Finally, we perform in the Preprocessor some program transformations tasks to simplify the analysis. For example, we convert all output commands $out(e)$ inside loops to an special assignment (Section~\ref{sec:program_trans1} and~\ref{sec:program_trans2}).

@@

\textbf{The Prover component} reduces the commutativity problem to a SMT formula satisfiability problem. At a high level, we check equivalence between a reducer program and its variant that has two consecutive inputs swapped. We show that this can be reduced to a SMT problem and give it to the SMT solver Z3~\cite{z3} for solving. In case that Z3 proved the formula unsatisfiable, we know that swapping any two consecutive inputs of the reducer will not change its output. Because all permutations of a list can be obtained by swapping consecutive elements finitely many times, it follows that the reducer outputs the same value for all permutations of the same list of inputs.
In this case,  J-ReCoVer stops and reports that the reducer is commutative. 
If we run the approach in a naive manner, it does not scale to large programs and is very imprecise.  We will introduce two critical optimizations that enables the proposed approach working for real-world examples in Section~\ref{section:optimizations}: a data flow analysis to extract critical variables and an improved symbolic execution algorithm to generate SMT formulae. More details of the algorithms used in the Preprocessor and the Prover components will be explained in Section~\ref{sec:preprocessor_prover} and the optimizations will be introduced in~\ref{section:optimizations}.

@@

Before entering the Prover component, the Preprocessor first performs a few program transformation tasks to simplify the verification task.
We assume the reducer $s_1;\rloop\{s_2\};s_3$ is given. The section begins with an introduction to the first program transformation task (Section~\ref{sec:program_trans1}) that handles the case when $s_2$ contains $out$ commands. The output of the first transformation task is a commutativity-equivalent reducer program where $out$ does not occur in $s_2$.  Then the algorithm of the Prover is introduced in Section~\ref{sec:prover}. This algorithm works only in a special case where no read from input occurs in $s_1$. For general reducer programs, we introduce the second program transformation task (Section~\ref{sec:program_trans2}) whose output is a reducer program that falls in the special case required by the Prover and ``over-approximates'' all input/output behaviors of the original reducer. The commutative of the produced reducer implies the commutative of the original one, but not vice verse.

@@

\subsection{The Prover: Reducing Commutativity to SMT Solving}
\label{sec:prover}
Below we begin with a simple version where $s_1$ never calls the $\cur$ function. Later we will explain how to handle the general case using program transformation. The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1} \rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$, where $n$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively. The construction of $F$ from $s_2$ can be done in the standard way. We will introduce a more efficient data structure for the construction in Section~\ref{section:optimizations}.
We check if the formula below is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$. 
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu}
\end{equation}

@@

In the example of ``average'', the values of $s$ and $c$ are the same no matter the reducer first reads $n_1$ and then $n_2$ or reads the two values in the swapped order.
So it computes the same value for the input [1;2;3;4;5] and any other input with two consecutive values swaps, e.g., [1;2;\textbf{4};\textbf{3};5].
Since any permutation of a sequence can be obtained by swapping consecutive values finitely many times~\cite{algebra}, this further implies that starting from the same variable valuation, the loop ends with the same valuation for any permutation of input values, e.g., the reducer will output the same value for [1;2;\textbf{3};\textbf{4};5], [1;\textbf{2};\textbf{4};3;5], [\textbf{1};\textbf{4};2;3;5], [4;1;2;3;5], etc. 
Recall that the first program transformation task remembers all output values in variables that will stay unchanged until the end of the loop. The validity of Equation~(\ref{eq:commu}) also implies that the execution of the loop produces the same sequence of outputs.

We explain why that the validity of Equation~(\ref{eq:commu}) is sufficient to prove the commutativity of the reducer $s_1;\rloop\{s_2\};s_3$. Notice that $s_1$ does not read from the input (assumed at the beginning of this subsection). The execution of $s_1$ is deterministic in the sense that from the same initial valuation, it outputs the same sequence of values and ends at the same valuation. For the execution of $\rloop\{s_2\}$, from the validity of Equation~(\ref{eq:commu}), any permutation of input values leads the reducer to the same final valuation and the same sequence of outputs at the end of the loop. 
Before executing the commands $s_3$, all inputs have been read (the loop exit condition) already. Similar the case of $s_1$, $s_3$ does not read from the input and hence the execution of $s_3$ is deterministic. It follows that for all permutations of inputs, the reducer outputs the same sequence of values.

In the next subsection, we explain how to generalize our approach to support general reducer programs using program transformation.

@@

We handle the problem again by program transformation. To handle the example in Figure~\ref{fig:reducer_max} (a), we move the prefix $s_1$ into the loop body and use a new variable $s$ to force that $s_1$ is always executed before the original loop body. The result after the transformation is demonstrated in Figure~\ref{fig:reducer_max} (b). The new reducer program over-approximates all possible inputs/outputs of the original one. For the case that $s=1$ at the beginning, its output sequence is the same as the original reducer. When $s\neq 1$, it may produce outputs that are not possible by the original reducer.
It also satisfies all assumptions required for the Prover component: 
(I) the commands before entering the loop does not have any $\cur$ expression, 
(II) each branch (recall that ``branch'' is formally defined in Section~\ref{section:integer-reducers}) of the loop body has at most one $\cur$, and 
(III) $out(v)$ command does not occur inside the loop body\footnote{We need to apply the first program transformation again to fulfill this assumption.}. So we can apply Equation~(\ref{eq:commu}) to prove its commutativity.
Since the new reducer has all the possible outputs of the original one, the new reducer being commutative implies that the original one is also commutative.

@@

We perform a simple backward data flow analysis to collect all variables whose value may propagate to the output command after the loop execution. With out loss of generality, we assume there is only one output command after the loop. Then the code after the loop (i.e., the $s_3$ part of the reducer) can only be in the form of Figure~\ref{fig:nested_out}.
We unfold all concatenation of commands, so now the commands $c_1\ldots c_n$ are either assignment commands $w:=e$ or branch command $\ite{g}{c'_1}{c'_2}$, and the only output command stays in the $m$-th layer of the branch command after $c_i$. For the case that $m=0$, the entire branch command in Lines 2-6 of Figure~\ref{fig:nested_out} is replaced with $out(v)$.

The data flow analysis algorithm works as follows.
A set $V_3$ whose value is initially a singleton set $\{v\}$ is used to remember important variables that might flow to the output command $out(v)$ in $s_3$ (Figure~\ref{fig:nested_out}).
The commands and guards will be inspected in the following order $c_{n}$,$\ldots$,$c_{i+1}$, $g_m$, $g_{m-1}$, $\ldots$, $g_1$, $c_i$, $\ldots$, $c_2$, $c_1$. For the case that a guard $g_j$ for $j\in [1,m]$ is being processed, we simply add all variables occurred in $g_j$ to $V_3$. 
For the case that a command $c_j$ for $j\in [1,n]$ is being processed, 
recall that $c_j$ can only be in the form of an assignment $w:=e$ or a branch $\ite{g}{c'_1}{c'_2}$.
For the case that $c_j$ is an assignment $w:=e$ and $w\in V_3$, we add all variables contained in $e$ to $V_3$. For the case that $c_j$ is a branch command $\ite{g}{c'_1}{c'_2}$, we process both $c'_1$ and $c'_2$ in a similar manner. Starting from the same set $V_3$,  after processing $c'_1$ and $c'_2$, we get two sets $V_3^1$ and $V_3^2$. We take their union $V_3^1\cup V_3^2$ as the new value of $V_3$. After the entire $s_3$ is being processed we get the set of important variables that might affect the outputs.

Then we modify the Equation~(\ref{eq:commu}) to the one below
\begin{equation}
F(n_1, F(n_2,x_1,x_2,\ldots,x_k) )=_{PosF(V_2\cup V_3)} F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu2}
\end{equation}
The function $PosF(X)$ returns the corresponding positions of variables in $X$ in the output of $F(n,x_1,\ldots,x_k) $. E.g., if $X = \{x_1,x_4,y\}$, where $y\notin \{x_1,\ldots,x_k\}$, then $PosF(X)=\{1,4\}$.
The use of equation~(\ref{eq:commu2}) is critical to the success of our approach. In our evaluation, the ratio of examples that our approach can successfully analyze is significantly increased from $6.8\%$ to $97.5\%$. The details of the evaluation can be found in Section~\ref{section:exp}.

@@

First, we define the set $\Var_0=\{v_0 \mid v \in \Var \}\cup \{c_0\}$ as the set of initial symbolic values of all variables, where $c_0$ is used to denote the symbolic value from the input.
A \emph{symbolic state} is a triple $\langle pc, val, cmds \rangle$, where the \emph{path condition} $pc$ is a guard in $\Grd$ over $\Var_0$, the \emph{symbolic valuation} is a mapping from variables in $\Var$ to an expression in $\Exp$ over $\Var_0$, and $cmds\in \Cmd$ is the remaining commands to be executed. The \emph{initial symbolic state} is $\langle \true, val_0, s_2\rangle$, where $val_0(v) = v_0$ for all $v\in\Var$ and $s_2$ is the loop body of the reducer. 
One can compute the next symbolic state of $\langle pc, val, c;cmds \rangle$ according to the following transition rules. Recall that $c$ is either in the form of $(v:= e_0)$ or $\ite{g}{c_1}{c_2}$ after the program transformation tasks are performed:
\begin{enumerate}
	\item If $c$ is $(v:= e_0)$, the next symbolic state is $\langle pc, val[v \rightarrow val(e_0) ], cmds \rangle$.
	\item If $c$ is $\ite{g}{c_1}{c_2}$, the next symbolic states are $\langle pc\wedge g, val, c_1.cmds \rangle$ and $\langle pc\wedge not\  g, val, c_2.cmds \rangle$. 
\end{enumerate}

Here we lift the $val$ function from variables to expressions in the standard way and $val[v \rightarrow val(e_0) ]$ is obtained by reassigning the value of $v$ in $val$ to $val(e_0)$. The execution from a symbolic state terminates when its path condition is conflict or the commands to be executed become empty, i.e., all commands have been processed. So far it is standard.

@@

We use a more efficient data structure to store symbolic values. At a high level, in a symbolic state, we count the number of occurrences of each symbolic values in $\Var_0$ and update the count when operations such as $+$ and $-$ is applied. For other operations $x\odot y$, we create a new symbolic expression $val(x) \odot val(y)$ and count its number of occurrence as one.
More concretely, we use a map $sv_x:\Exp\rightarrow \mathbb{N}$ (symbolic value of $x$) to remember the number of occurrences of each expression in $x$. We have $val(x)= (\Sigma_{z\in \dom{sv_x}} (z\times sv_x(z)) )$, where $\dom{sv}=\{e\mid sv(e)\neq 0\}$.
Initially, $sv_x(x_0) = 1$ and $sv_x (e) =0$ for all $e\neq x$. Assume that the symbolic value of $x,y$ are $sv_x,sv_y$ in the current symbolic state. The symbolic value $sv'_x$ of $x$ after the assignment $x:=x\circ y$ is defined as follows.

>>>>

MapReduce framework for data-parallel computation was first proposed by Google~\cite{dean04} and later implemented in the Apache Hadoop project~\cite{hadoop}.
Under the MapReduce framework, a reducer computes output values from a sequence of input values transmitted over the network.  Due to the non-determinism in data transmission, the order of input values arrives at the reducer is not fixed.
The \emph{commutativity problem} of reducers asks if the output of a reducer is independent of the order of its inputs. There are several advantages for a reducer being commutative, e.g., the verification problem of a MapReduce program can be reduced to the problem of verifying a sequential program. 
In this paper, we propose effective heuristics for reducer commutativity analysis and implement them as a tool J-ReCoVer (Java Reducer Commutativity Verifier), which is the first tool that is specialized in checking reducer commutativity. Experimental results over 128 benchmark examples collected from literature and open repositories are very positive; J-ReCoVer correctly handles over 97\% of them.

@@

On the other hand, in many cases, the non-commutative behavior of a reducer can be the source of some very tricky bugs. A study conducted by Microsoft investigated the commutativity problem of $508$ reducers running on their MapReduce server~\cite{xiao14mr}. The reducer programs have been checked carefully using all traditional means such as code review, testing, and experiments with real data for months. Still, among them, five programs contain very subtle bugs caused by non-commutativity (confirmed by the programmers).

@@

In this paper, we propose heuristics for the commutativity problem that works very efficiently on a large set of practical integer reducer programs. We reduce the commutativity problem to a SMT problem in a sound but incomplete manner. We also propose optimizations to enable the approach working on real world examples. For the case that the reducer is not proven commutative, we complement the approach by using testing to find concrete counterexamples. We implement all these heuristics as a tool called J-ReCoVer (Java Reducer Commutativity Verifier), which is available at~\verb|http://jrecover.iis.sinica.edu.tw|.

@@

One main motivation of commutativity analysis is to reduce the verification problem of a concurrent program to a sequential one. There are several researches sharing the similar motivation, e.g., the concept of \emph{robustness} of event-driven asynchronous programs~\cite{ahmed2017:robustness}, \emph{robustness} of program running under relaxed memory model~\cite{ahmed2013:robustness,AbdullaACLR13,AbdullaACLR12}, and \emph{synchronizability} of communicating finite state machines~\cite{FinkelL17,basu2012synchronizability,basu2011choreography}. However, their computation models are quite different to that of reducers and hence the result cannot be directly applied. Besides verification, another interesting research direction is the synthesis of MapReduce programs~\cite{SmithA16}. Commutativity analysis is used as a component under the synthesis framework.

\section{Notations and Definitions}
\label{section:integer-reducers}
We use $\Z$ to denote the set of all integers and $[n,m]$ to denote the set of integers $\{k\mid n \leq k\leq m\}$. 
We lift the equality $=$ operation to tuples in the standard way.
Let $S\subseteq [1,k]$, the equality $(v_1,\ldots,v_k)=_{S} (v'_1,\ldots,v'_k)$ between tuples holds if and only if $\bigwedge_{i\in S} v_i=v'_i$.
For presentation purpose, we describe a highly simplified model to characterize the essence of reducers. The model allows to describe the computation of reducers and investigate their commutativity.

\begin{equation*}
\begin{array}{rclcl}
\multicolumn{3}{l}{v\in \Var\mbox{, and }c \in \Z}&&\textmd{Variable and Constant}\\
e \in \Exp    &\ \ \  {\triangleq}\ \ \   & v \mid c \mid \cur \mid e + e \mid e - e \mid e \times e \mid *\mid\ldots & &\textmd{Integer Expression}\\
g\in \Grd  & {\triangleq}  & e < e \mid not\ g \mid g \wedge g  \mid \ldots& & \textmd{Guard}\\
s\in \Cmd  & {\triangleq}  & v:= e \mid out(v) \mid \ite{g}{s}{s} \mid s;s && \textmd{Command}\\
p\in \Prg  & {\triangleq}  & s;\rloop\{s\};s&& \textmd{Reducer Program}\\
\end{array}
\end{equation*}

Here $v\in \Var$ are variables and $c \in \Z$ are constant integer values, respectively. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{guard} in $\Grd$ can be a comparison between two expressions or boolean combinations of several guards. A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each \emph{branch}\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing every branch statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g); s_1$ or $(not\ g) ;s_2$. For example, the loop body of Figure~\ref{fig:reducer_example}(a) has two branches $t:=\cur; (t>0); m:=t$ and $t:=\cur;(not\ t>0)$ }. According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. A few examples of reducer programs in this model can be found in Figure~\ref{fig:reducer_example}.

@@

In the next section, we give a brief overview of J-ReCoVer,  our tool for reducer commutativity verification.

\section{Overview of the J-ReCoVer tool}
\label{sec:overview}
%Hadoop, Java;
The input of J-ReCoVer is a reducer program written in Java, which is the most popular programming language used in the Hadoop~\cite{hadoop} MapReduce framework. 
The J-ReCoVer tool has three main components, the \emph{Preprocessor}, the \emph{Prover}, and the \emph{BugFinder}. As the name suggests, the Preprocessor reads as input a reducer program and performs the required preprocessing.  The goal of Prover is to show that a given reducer is commutative and the goal of BugFinder is the opposite. The architecture of J-ReCoVer can be found in Figure~\ref{fig:overview-jrecover}. The user can input a reducer program to J-ReCoVer either through our web-interface or binary application installed on their own machine.

@@

\textbf{The Preprocessor component} first compiles a reducer program to bytecode and use the tool Soot~\cite{soot} to further convert it to the so called ``Jimple'' format, which is an intermediate language that is designed to simplify the analysis of Java programs. Under the Hadoop MapReduce framework, the permutation of the input is handled by the scheduler/shuffler component and is affected by things like network latency, which is not controllable by programmers. In order to deal with this issue, we wrote our own dummy Hadoop environment for the reducer as a part of the Preprocessor component so that the input order of the reducer is now controlled by J-ReCoVer. Finally, we perform in the Preprocessor some program transformation tasks to simplify the analysis. For example, we convert all output commands $out(e)$ inside loops to an special assignment (will be explained in Section~\ref{sec:program_trans1} and~\ref{sec:program_trans2}).

@@

\textbf{The Prover component} reduces the commutativity problem to a SMT formula satisfiability problem. At a high level, we check equivalence between a reducer program and its variant that has two consecutive inputs swapped. We show that this can be reduced to a SMT problem and give it to the SMT solver Z3~\cite{z3} for solving. In case that Z3 proved the formula unsatisfiable, we know that swapping any two consecutive inputs of the reducer will not change its output. Because all permutations of a list can be obtained by swapping consecutive elements finitely many times, it follows that the reducer outputs the same value for all permutations of the same list of inputs.
In this case,  J-ReCoVer stops and reports that the reducer is commutative. 
However, if we run the approach in a naive manner, it does not scale to large programs and is very imprecise.  We will introduce two critical optimizations that enables the proposed approach working for real-world examples in Section~\ref{section:optimizations}: a data flow analysis to extract critical variables and an improved symbolic execution algorithm to generate SMT formulae. More details of the algorithms used in the Preprocessor and the Prover components will be explained in Section~\ref{sec:preprocessor_prover} and the optimizations will be introduced in~\ref{section:optimizations}.

@@

Before entering the Prover component, the Preprocessor first performs a few program transformation tasks to simplify the verification task.
We assume the reducer $s_1;\rloop\{s_2\};s_3$ is given. The section begins with an introduction to the first program transformation task (Section~\ref{sec:program_trans1}) that handles the case when $s_2$ contains $out$ commands. The output of the first transformation task is a commutativity-equivalent reducer program where $out$ does not occur in $s_2$.  Then the algorithm of the Prover is explained in Section~\ref{sec:prover}. This algorithm works only in a special case where no read from input occurs in $s_1$. For general reducer programs, we introduce the second program transformation task (Section~\ref{sec:program_trans2}) whose output is a reducer program that falls in the special case required by the Prover and ``over-approximates'' all input/output behaviors of the original reducer. The commutative of the produced reducer implies the commutative of the original one, but not vice verse.

@@

\subsection{The Prover: Reduce Commutativity Checking to SMT Solving}
\label{sec:prover}
Below we begin with a simple reducer where $s_1$ never calls the $\cur$ function. Later we will explain how to handle the general case using program transformation. The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs the values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1} \rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$, where $n$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively, for $i\in[1,k]$. The construction of $F$ from $s_2$ can be done in the standard way. We will introduce a more efficient data structure for the construction in Section~\ref{section:optimizations}.
We check if the formula below is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$. 
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu}
\end{equation}

@@

In the example of ``average'', the values of $s$ and $c$ are the same no matter the reducer first reads $n_1$ and then $n_2$ or reads the two values in the swapped order.
So it computes the same value for the input [1;2;3;4;5] and any other input with two consecutive values being swapped, e.g., [1;2;\textbf{4};\textbf{3};5].
The swap operation can be applied arbitrarily many times to produce all possible permutations of the input. It follows that that starting from the same variable valuation, the loop ends with the same valuation for any permutation of input values, e.g., the reducer will output the same value for [1;2;\textbf{3};\textbf{4};5], [1;\textbf{2};\textbf{4};3;5], [\textbf{1};\textbf{4};2;3;5], [4;1;2;3;5], etc. 
Recall that the first program transformation task stores all output values in variables that will stay unchanged until the end of the loop. The validity of Equation~(\ref{eq:commu}) also implies that the execution of the loop produces the same sequence of outputs.

We explain why that the validity of Equation~(\ref{eq:commu}) is sufficient to prove the commutativity of the reducer $s_1;\rloop\{s_2\};s_3$. Notice that $s_1$ does not read from the input (assumed at the beginning of this subsection). The execution of $s_1$ is deterministic in the sense that from the same initial valuation, it outputs the same sequence of values and ends at the same valuation. For the execution of $\rloop\{s_2\}$, from the validity of Equation~(\ref{eq:commu}), any permutation of input values leads the reducer to the same final valuation and the same sequence of outputs at the end of the loop. 
Before executing the commands $s_3$, all inputs have been read (the loop exit condition) already. Similar to the case of $s_1$, $s_3$ does not read from the input and hence the execution of $s_3$ is deterministic. It follows that for all permutations of inputs, the reducer outputs the same sequence of values.

In the next part, we explain how to generalize our approach to support general reducer programs (those without the restriction on $s_1$) using program transformation.

@@

We handle the problem again by program transformation. To deal with the example in Figure~\ref{fig:reducer_max} (a), we move the prefix $s_1$ into the loop body and use a new variable $s$ to force that the execution of $s_1$ is always before the original loop body. The result after the transformation is demonstrated in Figure~\ref{fig:reducer_max} (b). The new reducer program over-approximates all possible inputs/outputs of the original one. For the case that $s=1$ at the beginning, its output sequence is the same as the original reducer. When $s\neq 1$, it may produce outputs that are not possible by the original reducer.
It also satisfies all assumptions required for the Prover component: 
(I) the commands before entering the loop does not have any $\cur$ expression, 
(II) each branch (recall that ``branch'' is formally defined in Section~\ref{section:integer-reducers}) of the loop body has at most one $\cur$, and 
(III) $out(v)$ command does not occur inside the loop body\footnote{We need to apply the first program transformation again to fulfill this assumption.}. So we can apply Equation~(\ref{eq:commu}) to prove its commutativity.
Since the new reducer has all the possible outputs of the original one, the new reducer being commutative implies that the original one is also commutative.

@@

We perform a simple backward data flow analysis to collect all variables whose value may propagate to the output command after the loop execution. Without loss of generality, we assume there is only one output command after the loop. Then the code after the loop (i.e., the $s_3$ part of the reducer) can only be in the form of Figure~\ref{fig:nested_out}.
We unfold all concatenation of commands, so now the commands $c_1\ldots c_n$ are either assignment commands $w:=e$ or branch command $\ite{g}{c'_1}{c'_2}$, and the only output command stays in the $m$-th layer of the branch command after $c_i$. For the case that $m=0$, the entire branch command in Lines 2-6 of Figure~\ref{fig:nested_out} is replaced with $out(v)$.

The data flow analysis algorithm works as follows.
A set $V_3$ whose value is initially a singleton set $\{v\}$ is used to remember important variables that might flow to the output command $out(v)$ in $s_3$ (in Figure~\ref{fig:nested_out}).
The commands and guards will be inspected in the following order $c_{n}$,$\ldots$,$c_{i+1}$, $g_m$, $g_{m-1}$, $\ldots$, $g_1$, $c_i$, $\ldots$, $c_2$, $c_1$. For the case that a guard $g_j$ for $j\in [1,m]$ is being processed, we simply add all variables occurred in $g_j$ to $V_3$. 
For the case that a command $c_j$ for $j\in [1,n]$ is being processed, we need to handle two different cases:
\begin{itemize}
	\item $c_j$ is an assignment $w:=e$ and $w\in V_3$: all variables in $e$ are added to $V_3$. 
	\item $c_j$ is a branch $\ite{g}{c'_1}{c'_2}$: the algorithm starts from the same set $V_3$ and processes $c'_1$ and $c'_2$ individually and independently (in a similar manner) to obtain $V_3^1$ and $V_3^2$. After that, the algorithm takes their union $V_3^1\cup V_3^2$ as the new value of $V_3$.
\end{itemize}
After the entire $s_3$ is being processed we get the set of important variables that might affect the outputs that happen in $s_3$.
Then we modify the Equation~(\ref{eq:commu}) to the one below
\begin{equation}
F(n_1, F(n_2,x_1,x_2,\ldots,x_k) )=_{PosF(V_2\cup V_3)} F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu2}
\end{equation}
The function $PosF(X)$ returns the corresponding positions of variables in $X$ in the output of $F(n,x_1,\ldots,x_k) $. E.g., if $X = \{x_1,x_4,y\}$, where $y\notin \{x_1,\ldots,x_k\}$, then $PosF(X)=\{1,4\}$.
The use of equation~(\ref{eq:commu2}) is critical to the success of our approach. In our evaluation, the ratio of examples that our approach can successfully analyze is significantly increased from $6.8\%$ to $97.5\%$. The details of the evaluation can be found in Section~\ref{section:exp}.

@@

First, we define the set $\Var_0=\{v_0 \mid v \in \Var \}\cup \{c_0\}$ as the set of initial symbolic values of all variables, where $c_0$ is used to denote the symbolic value from the input.
A \emph{symbolic state} is a triple $\langle pc, val, cmds \rangle$, where the \emph{path condition} $pc$ is a guard in $\Grd$ over $\Var_0$, the \emph{symbolic valuation} is a mapping from variables in $\Var$ to an expression in $\Exp$ over $\Var_0$, and $cmds\in \Cmd$ is the remaining commands to be executed. The \emph{initial symbolic state} is $\langle \true, val_0, s_2\rangle$, where $val_0(v) = v_0$ for all $v\in\Var$ and $s_2$ is the loop body of the reducer. 
One can compute the next symbolic state of $\langle pc, val, c;cmds \rangle$ according to the following transition rules. Recall that $c$ is either in the form of $(v:= e)$ or $\ite{g}{c_1}{c_2}$ after the program transformation tasks are performed:
\begin{enumerate}
	\item If $c$ is $(v:= e)$, the next symbolic state is $\langle pc, val[v \rightarrow val(e) ], cmds \rangle$.
	\item If $c$ is $\ite{g}{c_1}{c_2}$, the next symbolic states are $\langle pc\wedge g, val, c_1.cmds \rangle$ and $\langle pc\wedge not\  g, val, c_2.cmds \rangle$. 
\end{enumerate}

Here we lift the $val$ function from variables to expressions in the standard way and $val[v \rightarrow val(e) ]$ is obtained by reassigning the value of $v$ in $val$ to $val(e)$. The execution from a symbolic state terminates when its path condition is conflict or the commands to be executed become empty, i.e., all commands have been processed. So far it is standard.

@@

We use a more efficient data structure to store symbolic values. At a high level, in a symbolic state, we count the number of occurrences of each symbolic values in $\Var_0$ and update the count when operations such as $+$ and $-$ is applied. For other operations $x\odot y$, we create a new symbolic expression $val(x) \odot val(y)$ and count its number of occurrence as one.
More concretely, we use a map $sv_x:\Exp\rightarrow \mathbb{N}$ (symbolic value of $x$) to remember the number of occurrences of each expression in $x$. We have $val(x)= (\Sigma_{z\in \dom{sv_x}} (z\times sv_x(z)) )$, where $\dom{sv}=\{e\mid sv(e)\neq 0\}$.
Initially, $sv_x(x_0) = 1$ and $sv_x (e) =0$ for all $e\neq x$. Assume that the symbolic value of $x,y$ are $sv_x,sv_y$ in the current symbolic state. The symbolic value $sv'_x$ of $x$ after the assignment $x:=p\circ q$ is defined as follows.
